### 2019-01-20 tex_v2/cover_letter.tex
rewrite and restructure narrative

\documentclass[10pt,letterpaper]{letter}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[letterpaper,left=1in, right=1in, top=1in, bottom=1in]{geometry}

\makeatletter
\let\@texttop\relax
\makeatother

\address{Andrew Jansen\\
Life Sciences A-Wing\\
Arizona State University\\
427 E. Tyler Mall\\
PO Box 874501\\
Tempe, AZ 85287-4501} 
\signature{Andrew Jansen\\
Corresponding Author – majanse1@asu.edu\\
Graduate Student\\
School of Life Sciences\\
Arizona State University\\
Tempe, AZ 85287-4501, USA} 
\begin{document} 
\begin{letter}{NAME\\
Editor\\
email@address\\
address 1\\
address 2\\
address 3}
\opening{Dear. !!!EDITOR-NAME-HERE!!!,}

I am writing to submit our manuscript entitled, ``Microstructure-derived strength in the acorn weevil exoskeleton'', for consideration for publication in Nature Materials.
This paper DOES THIS NEW IMPACTFUL THING.

We WHAT WE DO HEREIN.
This work represents WHAT THIS WORK SIGNIFIES.

Our manuscript therefore serves two distinct, but equally important ends: (1) PURPOSE 1, and (2) PURPOSE 2.
All of the authors have agreed to submit the manuscript to Nature Materials; we feel that this journal will permit us to reach the ideal readership to disseminate our findings and to promote these methods.

We suggest the following individuals be contacted as potential reviewers: Dr. ONE (EMAIL), for WHAT REASON, as well as Dr. TWO (EMAIL) and Dr. THREE (email), both of whom are REASON.

This paper is not being considered for publication at any other journal.
We have carefully followed your manuscript preparation guidelines, and look forward to your comments.
If this manuscript is not suitable to your needs, please notify us as soon as possible.
We thank you for your time and consideration.

\closing{Sincerely,}

\newpage

\cc{Jason Williams\\
Co-author\\
Research Associate\\
School for Engineering of Matter, Transport, and Energy\\
Arizona State University\\
Tempe, AZ 85287-4501, USA\\
\\
Nikhilesh Chawla\\
Co-author\\
Professor\\
School for Engineering of Matter, Transport, and Energy\\
Arizona State University\\
Tempe, AZ 85287-4501, USA\\
\\
Nico Franz\\
Co-author\\
Professor\\
School of Life Sciences\\
Arizona State University\\
Tempe, AZ 85287-4501, USA} 
%\ps{adding a postscript} 
%\encl{list of enclosed material} 
\end{letter} 
\end{document}